---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

knitr::opts_chunk$set(echo = TRUE)

## Library useful packages
```{r}
library(ggplot2)
library(tidyverse)
```

## Loading and preprocessing the data
```{r}
activity <- read.table(unz("activity.zip", "activity.csv"), header = TRUE, sep=",", stringsAsFactors = FALSE)
```



## What is mean total number of steps taken per day?
1. Calculate total number of steps 
```{r}
StepsByday <- activity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))
```

2. Make the Histogram for total number of steps taken per da  y
```{r}
g <- ggplot(StepsByday, aes(x = total))
g + geom_histogram(binwidth = 3000, colour = "#000000") + labs(title = "Steps taken per day", x = "Steps", y = "Frequency")
```

3. Calculate the mean total number of steps taken per day
```{r}
mean_steps <- mean(StepsByday$total)
median_steps <- median(StepsByday$total)
mean_steps
median_steps
```

## What is the average daily activity pattern?
1.Time series plot of the average number of steps taken
```{r}
StepsByIntervals <- activity %>% group_by(interval) %>% summarise(mean = mean(steps, na.rm = TRUE))
g <- ggplot(StepsByIntervals, aes(x = interval, y = mean))
g + geom_line() + labs(title = "Average number of steps taken by 5 min intervals", x = "Intervals(5min)", y = "Average steps taken")
```

2.The 5-minute interval that, on average, contains the maximum number of steps
```{r}
maxsteps <- StepsByIntervals[which.max(StepsByIntervals$mean),1]
maxsteps
```

## Imputing missing values
1.  Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(activity$steps))
```

2.Impute missing value, using daily average steps divided by total intervals per day.
```{r}
activity_new <- activity
activity_new$steps[is.na(activity_new$steps)] <-mean_steps/288 
```

3.Histogram of the total number of steps taken each day after missing values are imputed
```{r}
StepsByday_r <- activity_new %>% group_by(date) %>% summarise(total = sum(steps)) 
g <- ggplot(StepsByday_r, aes(total))
g + geom_histogram(binwidth = 3000, colour = "#000000") + labs(title = "Total number of steps taken each day", x = "Steps", y = "Frequency") 
```
4. Calculate the Mean and Median after imputing missing value
```{r}
mean_steps <- mean(StepsByday$total)
median_steps <- median(StepsByday$total)
mean_steps
median_steps
```
## Are there differences in activity patterns between weekdays and weekends?
1. Define weekdays and weekends
```{r}
activity_new$dateType <-  ifelse(as.POSIXlt(activity_new$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

2. Histogram

```{r}
activity_weekcom <- activity_new %>% group_by(interval, dateType) %>% summarise(mean = mean(steps))
g <- ggplot(activity_weekcom, aes(interval, mean))
g + geom_line(aes(col = dateType)) + labs(title = "Total number of steps taken per day", x = "Interval", y = "Average steps") + scale_colour_discrete(name  = "Date Type")
```
